Impact of frequent assimilation of surface and satellite observations in an MCS case during RELAMPAGO field campaign. Part I: impact on the analysis

### 2021-07-26 analysis/chapters/paper_ana_20181122.Rmd
Revisión del texto

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing it with each type of observations. $sd RCRV$ for  wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error. For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas to the west of the domain where the $sd RCRV$ can be as high as 4.5. In this region the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.

Figure \@ref(fig:rcrv-profile) shows the mean and standard deviation of the RCRV for the upper-air observations. Figure \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error. As sounding observations and the associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with values that suggests that the error for this type of observation is overestimated. ADPUPA and AIRCAR also present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dillonInreview.

For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behaviour with a negative bias and an overestimation of the observation error. Wind estimations derived from water vapor channels are abundant over 500 hPa  where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) shows almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor indicating that the bias correction algorithm works as expected. The $sd RCRV$ values are less than 1 for all sensors possibly due to an overestimation of the observation errors to reduce the influence of potentially erroneous observations.

Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work. Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate.

@@

In this section we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection. The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS. Figure \@ref(fig:TQ-diff)a-c shows the difference between AUT and CONV in the spatially averaged vertical profile of temperature. By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state. During the first day, the assimilation of ASWS observations results in a colder PBL. This cooling effect has a clear diurnal cycle being stronger during night time. (Figure \@ref(fig:TQ-diff)a). During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS. The warm difference  shown in AUT with respect to CONV  in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV. This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection. Figure \@ref(fig:TQ-diff)b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment. This result is unexpected since mid-to-upper level winds can have a profound impact in the organization and evolution of MCSs over the region. 
During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensate for the cooling effect of ASWS observations. No clear systematic impact is found above the PBL during this period. During the second day, the impact of radiance observations is found through the troposphere with a distribution which is similar to the impact found in the AUT experiment but with opposite sign.

Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \@ref(fig:TQ-diff)d). The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the night time. The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010]. We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL. In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections. As for the temperature, the systematic impact of satellite-derived winds is smaller and only important after 18 UTC Nov, 21 at low levels. Figure \@ref(fig:TQ-diff)f shows that, during the first day of the experiment, radiances reduce low-middle level moisture. The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.

@@

The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis. There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain. During the second day of the experiment the systematic impact on the wind field is larger at mid and upper levels. During Nov 22 and 23 the impact of assimilating ASWS, produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however the systematic impact is overall smaller than the one produced by assimilating data from ASWS.

The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure  \@ref(fig:UV-diff)c and f). For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of ASWS observations. On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind. When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind.

To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles). At that moment the first convective cells were developing on the southern region of the domain along the cold front. Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AUT and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AUT and SATWND but not higher than CONV. The distribution of moisture at low levels in RAD seems to be the results of the combination of the moistening effect of assimilating ASWS, partially compensated by the assimilation of radiance observations, and a reduced meridional moisture transport due to the weaker northerly flow (with respect to CONV) over the centre of the domain.

The potential temperature on the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed on the temperature profiles in Figure \@ref(fig:TQ-diff)a-c. On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations. Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear. The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$. MCAPE in the RAD experiment is lower compared to CONV or SATWND. This is consistent with less humidity in the PBL with respect to these experiments and maybe partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AUT, SATWND and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values. Moreover in this same region, these experiments show larger MCAPE values. Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells (e.g. Markowsky and Richardson 2010).

@@

Firstly we analyse the impact of different observation types in terms of the representation of the MCS and its associated precipitation. Figure \@ref(fig_pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and estimated precipitation by IMERG. The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the experiments. In all the experiments, the accumulated precipitation in the short-range forecasts underestimate the precipitation. This is particularly evident on CONV where the convection initiation is significantly delayed with respect to the observations and occurs further north with respect to the observed initiation. AUT, SATWND and RAD better capture the timing and location of convective initiation. On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue. After 18 UTC Nov, 22 RAD shows improvements in precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the conveccion.

@@

To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss). All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature and dissipation stages of the MCS. The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the maturity and dissipation stages of the MCS. AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study. The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22. The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).

@@

Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes). IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments. The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection. For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that are slightly reduced by the assimilation of ASWS and radiances. The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND and RAD is small above the PBL. The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND and RAD. On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C), indicating a more moist atmosphere in the analysis. Overall, the RMSE grows with height, this is probably because at upper levels, moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures. The zonal wind has a significant positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced by assimilating satellite radiances. It is somewhat surprising that no significant effect is observed by assimilating satellite derived winds. The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.

@@

For the IOP 8 (bottom row in Figure \@ref(fig:soundings)), the densely observed area was behind the MCS, but far enough from it to be directly affected by its mesoscale circulations. This area was also behind the cold front thus affected by low-level cold advection. Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances. Between 5 and 12 km, the bias in each experiment is cold. AUT, SATWND and RAD show a lower bias due to the assimilations of ASWS. The RMSE for AUT, SATWND and RAD remains below 2$^{\circ}$C for almost all levels with significant  improvement in upper levels where radiance observations have the largest impact. In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations. This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region. The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.  SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C. The zonal wind is overestimated on the analyses with a positive bias at middle levels. For this period only RAD shows an improvement in the upper troposphere. At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND and RAD. In fact low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).

>>>>

To investigate the ability of the first-guess ensemble mean to fit the observations taking into account the uncertainties of the forecast and the observations, we calculated the $mean RCRV$ and the $sd RCRV$ for the RAD experiment. As this experiment assimilates all types of observations used in this work, it is possible to analyse the reliability of the ensemble by comparing it with each type of observations. Figure 5 shows the $sd RCRV$ for surface observations box-averaged to a 2.5° grid. The $sd RCRV$ for wind observations (Figure \@ref(fig:rcrv-sfc)a) is close to 1 suggesting a good agreement between the ensemble spread, the forecast error and the observation error. For the temperature (Figure \@ref(fig:rcrv-sfc)b), the results are similar except that for some areas to the west of the domain where the $sd RCRV$ can be as high as 4.5. In this region the mean innovation was overly large due to the complex terrain leading to large differences between the model topography and the station height.

Figure \@ref(fig:rcrv-profile) shows the mean and standard deviation of the RCRV for the upper-air observations. Figure \@ref(fig:rcrv-profile)a-b show the RCRV statistics for soundings (ADPUPA) and aircraft (AIRCAR and AIRCFT). Both ADPUPA and AIRCFR show a general good agreement between the ensemble spread and the observation error. As sounding observations and their associated errors are known to be reliable, this result indicates that the ensemble has an appropriate spread. AIRCAR presents an irregular profile with $sd RCRV$ values that suggests that the error for this type of observation is overestimated. ADPUPA and AIRCAR present a $mean RCRV$ profile near zero on middle and upper levels but ADPUPA observations show a cold bias respect to the model at low levels, a characteristic already studied in @ruiz2010 and @dilloninreview.

For satellite-derived winds (Figure \@ref(fig:rcrv-profile)c) in low levels, where the are not many observations available, the profiles of $mean RCRV$ and $sd RCRV$ shows a larger departure from the expected behaviour with a negative bias and an overestimation of the observation error. Wind estimations derived from water vapor channels are abundant over 500 hPa  where their bias is close to zero. The only exception are the EUMETSAT observations which contribute very little in the region.

The mean RCRV profiles calculated from the radiance observations (Figure \@ref(fig:rcrv-profile)d) shows almost no bias and the same happens if the $mean RCRV$ is calculated over each channel of each sensor. This indicates that the bias correction algorithm works as expected. The $sd RCRV$ values are less than 1 for all sensors possibly due to an overestimation of the observation errors to reduce the influence of potentially erroneous observations.

Overall these results indicate that the ensemble spread is consistent with the short-range forecast error and that systematic errors are relatively small for most of the observation types used in this work. Moreover, these results suggest the inflation parameter $\beta = 0.9$ is adequate for the system.

@@

In this section we present the impact of assimilating different observation types on variables which are particularly relevant for the occurrence of deep moist convection. The analysis is performed over a smaller domain (red box in Figure \@ref(fig:dominio)a) to focus on the region most directly affected by the MCS. Figure \@ref(fig:TQ-diff)a-c shows the difference between experiments in the spatially averaged vertical profile of temperature. By averaging the differences between two experiments we can isolate the systematic impact produced by different observing systems on the analyzed state. During the first day, the assimilation of ASWS observations results in a colder PBL. This cooling effect has a clear diurnal cycle being stronger during night time (Figure \@ref(fig:TQ-diff)a). During the second day of the experiment, the impact of ASWS observations extends into the middle and upper troposphere coinciding with the mature stage of the MCS. The warm difference  shown in AUT with respect to CONV  in mid and upper levels is produced by the development of stronger convection in this experiment compared to CONV. This is a good example of how low-level information provided by surface weather stations can rapidly spread into the troposphere in the presence of deep moist convection. Figure \@ref(fig:TQ-diff)b, indicates that SATWND observations have an almost negligible impact on the mean temperature through the experiment. This result is unexpected since mid-to-upper level winds can have a profound impact in the organization and evolution of MCSs over the region. 
During the first day of the experiment, the assimilation of radiances produces a warming effect in the PBL which partially compensate for the cooling effect of ASWS observations. No clear systematic impact is found above the PBL during this period. During the second day, the impact of radiance observations is found through the troposphere with a distribution that is similar to the impact found in the AUT experiment but with opposite sign.

Comparing the specific humidity in the experiments (Figure \@ref(fig:TQ-diff)d-f) the impact of assimilating ASWS with larger spatial and temporal resolution is most significant at low levels (Figure \@ref(fig:TQ-diff)d). The PBL in the AUT experiment is consistently moister than in the CONV experiment and this is enhanced during the night time. The increase in low-level moisture by a denser surface network is consistent with previously reported dry biases in the WRF model over the region [@ruiz2010]. We found that the moistening of the PBL is mainly driven by the covariance between temperature and specific humidity within the PBL. In our experiment and over the center of the domain, this covariance remains negative, increasing low-level moisture as the observations introduce negative temperature corrections. As for the temperature, the systematic impact of satellite-derived winds is small and only important after 18 UTC Nov, 21 at low levels. Figure \@ref(fig:TQ-diff)f shows that radiances reduce low-middle level moisture during the first day of the experiment. The drying effect extends to lower-middle levels during the second day of the experiment coinciding with the development of the MCS between 00 and 12 UTC Nov, 22.

@@

The assimilation of ASWS produces an increase of zonal wind and a decrease of meridional wind at low levels during the first two days of analysis. There is a diurnal cycle in the impact of surface weather stations on the meridional velocity  (Figure  \@ref(fig:UV-diff)d) with stronger reduction of the northerly wind during night hours. This indicates that surface observations are reducing the intensity of the low level jet present in the pre-convective environment. After 18 UTC Nov, 22 the opposite effect is observed when the MCS is moving through the domain. After the initiation of the convective cells the systematic impact on the wind field is larger at mid and upper levels. During Nov 22 and 23 the impact of assimilating ASWS, produces an increase of northerly wind in upper levels. This could be a consequence of a stronger MCS with an increased polar side upper level outflow. Satellite-derived wind observations produce the largest impact in mid-to-upper levels where the number of observations is larger, however the systematic impact is overall smaller than the one produced by assimilating data from ASWS.

The assimilation of radiances produced a reduction in the westerly wind compared with respect to SATWIND in low and upper levels (Figure  \@ref(fig:UV-diff)c and f). For the meridional wind, these observations produce an enhancement on average of the northerly low-level flow of $1 ms^{-1}$, opposite to what is generated by the assimilation of ASWS observations. On upper levels and during Nov 22 and 23 the average impact of assimilating radiances is a decrease in the wind. When we look at the meridional wind field at 200 hPa at different times, we observed that the outflow from the MCS is even more intense than in the other experiments while the southerly wind ahead of the MCS also increases producing in average a reduction of the northerly wind present in  Figure /@ref(fig:UV-diff)f.

To investigate how changes in the PBL can modify the pre-convective environment we compare the horizontal distribution of the low level northerly flow, precipitable water, low level temperature and CAPE for the analysis at 00 UTC Nov, 22 (after 30 assimilation cycles). At that moment the first convective cells were developing on the southern region of the domain along the cold front. Figure \@ref(fig:summary-fields)a shows the precipitable water (shaded) and the vertically averaged low-level meridional wind component (contours). We found that the moist tongue extending over the northern part of the domain is enhanced by the assimilation of denser surface observations. The moisture increase is particularly stronger at the southern tip of this tongue, just ahead of the cold front where convection initiation was taking place. AUT and SATWND experiments are very similar, with values of precipitable water over 55 $kgm^{-2}$ north of 30$^{\circ}$S and a similar vertical distribution of specific humidity (not shown). RAD has lower precipitable water content than AUT and SATWND but higher than CONV. The distribution of moisture at low levels in RAD seems to be the results of the combination of the moistening effect of assimilating ASWS  -- partially compensated by the assimilation of radiance observations -- and a reduced meridional moisture transport due to the weaker northerly flow over the centre of the domain compare to CONV.

The potential temperature on the PBL (Figure \@ref(fig:summary-fields)b) resembles the characteristics observed on the temperature profiles in Figure \@ref(fig:TQ-diff)a-c. On average the PBL in AUT and SATWND is colder than in CONV while RAD shows a warmer PBL than AUT due to the assimilation of radiance observations. Figure \@ref(fig:summary-fields)c shows the most unstable convective available potential energy (MCAPE, shaded) and the 0 to 6 $km$ wind shear. The values of MCAPE in CONV do not exceed 2000 $JKg^{-1}$ while the rest of the experiments show maximum MCAPE over 4000 $JKg^{-1}$. MCAPE in the RAD experiment is lower compared to AUT or SATWND. This is consistent with less humidity in the PBL with respect to these experiments and maybe partially compensated by a slightly warmer PBL in the RAD experiment. The 0-6 km wind shear is more intense in AUT, SATWND and RAD reaching values over 15 $ms^{-1}$ at the southern tip of the region with positive MCAPE values. Moreover in this same region, these experiments show larger MCAPE values. Wind shear over 15 $ms^{-1}$ have been associated with the development of more intense and organized MCSs [@chen2015] and also with conditions favorable for supercells [@markowsky2010].

@@

Firstly we analyse the impact of different observation types in terms of the representation of the MCS and its associated precipitation. Figure \@ref(fig:pp-hov) shows the first-guess hourly accumulated precipitation averaged over a longitud range between 67$^{\circ}$W and 54.5$^{\circ}$W as a function of time and latitude in the different experiments, and the estimated precipitation by IMERG. The heaviest precipitation (over 12 $mmh^{-1}$) starts during the afternoon of Nov 22 and continues during Nov 23 after the end of the simulated period. In all the experiments, the accumulated precipitation in the short-range forecasts underestimate the precipitation. This is particularly evident on CONV where the convection initiation is significantly delayed and occurs further north with respect to the observed initiation. AUT, SATWND and RAD better capture the timing and location of convective initiation. On RAD the initiation is delayed with respect to AUT and SATWND possibly due to the lower content of precipitable water and MCAPE on the southern tip of the moist tongue. After 18 UTC Nov, 22 RAD shows improvements in the precipitation rate and its distribution compared to the other experiments as a result of the enhanced development of the conveccion.

@@

To quantify the spatial match between the observed precipitation and the short range forecasted precipitation for the different experiments, we compute the FSS in 6-hours moving windows for different thresholds and spatial scales (Figure \@ref(fig:fss)). All experiments show low values of FSS before and during the initiation of the MCS around 03 UTC Nov, 22, with an improvement during the mature stage of the MCS. The FSS for CONV is the lowest compared to the rest of the experiments and the differences are larger during the mature stage of the MCS. AUT and SATWND show similar FSSs indicating that satellite-derived wind assimilation has little impact on the precipitation for the present case study. The assimilation of radiances led to an overall improvement of the 1-hour forecasted precipitation, particularly for the 25 mm threshold during the period of heaviest precipitation on Nov, 22. The enhancement is also important at the developing stage of the MCS (between 00 and 12 UTC Nov, 22 and for spatial scales above 500 km, not shown).

@@

Finally, Figure \@ref(fig:soundings) shows the RMSE and bias calculated by comparing the experiments with radiosonde data from the RELAMPAGO missions, IOP 7 from 15 to 21 UTC Nov, 21 (including 30 radiosondes) and IOP 8 from 14 to 20 UTC Nov 22 (including 22 radiosondes). IOP 7 provides a good characterization of the pre-convective environment during the first day of our experiments. The area where the observations were taken was characterized by mostly clear skies and a low-level northerly flow associated with warm and moist advection. For the temperature, CONV presents a warm bias on the PBL of 1$^{\circ}$C and an RMSE between 1 and 2.5$^{\circ}$C that are slightly reduced by the assimilation of ASWS and radiances. The RMSE for temperature remains between 0.5 and 2$^{\circ}$C throughout the troposphere and lower stratosphere and the impact of additional observations in the AUT, SATWND and RAD is small above the PBL. The dew point temperature shows a small dry bias in low levels that is corrected on AUT, SATWND and RAD. On middle levels (below 10 km) the bias is large and positive (8 - 10$^{\circ}$C), indicating a more moist atmosphere in the analysis. Overall, the RMSE grows with height, this is probably because at upper levels, moisture contents are lower and small changes in the amount of water vapor can produce relatively large differences in terms of dew point temperatures. The zonal wind has a significant positive bias of approximately 10 $ms^{-1}$ in the upper troposphere which is mainly reduced by assimilating satellite radiances. It is somewhat surprising that no significant effect is observed by assimilating satellite derived winds. The assimilation of ASWS and radiance observations increases the northerly wind bias and RMSE near the surface.

@@

For the IOP 8 (bottom row in Figure \@ref(fig:soundings)), the densely observed area was behind the MCS, but far enough from it to be directly affected by its mesoscale circulations. This area was also behind the cold front thus affected by low-level cold advection. Below 5km all experiments show a warm bias which decreases with the assimilation of ASWS and radiances. Between 5 and 12 km, the bias in each experiment is cold and is reduced in AUT, SATWND and RAD to the assimilations of ASWS. The RMSE for AUT, SATWND and RAD remains below 2$^{\circ}$C for almost all levels with significant  improvement in upper levels where radiance observations have the largest impact. In terms of the dew point temperature, the analyses have a moist bias near the surface that increases with the incorporation of new observations. This moist bias is mostly driven by the assimilation of surface temperature observations and the fact that near the surface there is a negative covariance between temperature and moisture content over this region. The reason why such a covariance structure emerges from the ensemble is not clear and deserves further investigation but it is evident that the corrections introduced by the observations are degrading the estimation of the moisture content within the PBL.  SATWND presents the lowest RMSE overall while AUT and RAD are comparable between 4 and 12 km with values that do not exceed 7$^{\circ}$C. The zonal wind is overestimated on the analyses with a positive bias at middle levels. For this period only RAD shows an improvement in the upper troposphere. At low levels the meridional wind presents a negative bias, indicating an underestimation of the southerly wind behind the cold front principally in AUT, SATWND and RAD. In fact low level biases, in these experiments, are higher than in the CONV experiment indicating a detrimental effect of the additional observations (possibly associated with the effect of ASWS).
